---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# visualizeNOAA

<!-- badges: start -->
<!-- badges: end -->

# Introduction
  
This vignette provides an overview of the `visualizeNOAA` package functionalities for cleaning and visualizing earthquake data from NOAA.

## Installation

You can install the package from GitHub using the following command:
```{r setup}
#devtools::install_github("apd/visualizeNOAA")
```

## Loading the Package
Once installed, load the package into your R session:
```{r}
library(visualizeNOAA)
library(tidyverse)
library(leaflet)
```


## Cleaning NOAA Earthquake Data
The eq_clean_data() function cleans the raw NOAA earthquake data. It reads the data from a tab-delimited file, creates a date column, and converts latitude and longitude to numeric class.
```{r}
cleaned_data <- eq_clean_data("inst/extdata/earthquakes.tsv")
cleaned_data %>%
  head()
```

## Cleaning Earthquake Location Names
The eq_location_clean() function cleans the LOCATION_NAME column by stripping out the country name and converting names to title case.
```{r}
cleaned_location <- eq_location_clean(cleaned_data)
cleaned_location %>% 
  head()
```

## Visualizing Earthquake Data
### Timeline of Earthquakes
The geom_timeline() function plots a timeline of earthquakes.
```{r, fig.dim = c(6, 4)}
df_visualize <- bind_rows(
  cleaned_location %>%
    head(10),
  data.frame(country="", location ="")) %>% 
  mutate(country=factor(country, levels = unique(country))) 
ggplot(data = df_visualize) +
  geom_timeline(aes(date,size = magnitude,  color = no_deaths)) +
  theme_minimal() +
  theme(
    axis.line.x = element_line(linewidth = 0.75),
    axis.title.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.position = "bottom"
  )

```

### Timeline of Earthquakes
The geom_timelinelabel() function includes label as well.
```{r, fig.dim = c(6, 4)}
ggplot(data = df_visualize) +
  geom_timeline(aes(date,country, size = magnitude,  color = no_deaths))  +
  geom_timelinelabel(aes(date,country,label = location), n_max = 5) +
  theme_minimal() +
  theme(
    axis.line.x = element_line(linewidth = 0.75),
    axis.title.y = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    legend.position = "bottom"
  )
```

